---
title: "Lab 05 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Shatavia Bellmon"
date: "2/23/26"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(readr) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
dennys <- read_csv("data/dennys.csv")
laquinta <- read_csv("data/laquinta.csv")
```

### Exercise 1

```{r}
dn <- dennys
```

```{r}
dn_ak <- dn %>%
  filter(state == "AK")
nrow(dn_ak)
```

There are three Denny's in Alaska.

```{r}
lq <- laquinta
```

```{r}
lq_ak <- lq %>%
  filter(state == "AK")
nrow(lq_ak)
```

There are two Laquintas in Alaska.

### Exercise 2

```{r}
num_parings <- nrow(dn_ak) * nrow((lq_ak))
```

There needs to be 6 parings.

### Exercise 3

```{r}
dn_lq_ak <- full_join(dn_ak, lq_ak,
  by = "state"
)
dn_lq_ak
```


### Exercise 4

View(dn_lq_ak)

There are six observations in the dn_lq_ak data frame. The variables in this data frame are address.x, city.x, state, zip.x, longitude.x, latitude.x, address.y, city.y, zip.y, longitude.y, and latitude.y.

### Exercise 5

```{r}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 <- long1 * pi / 180
  lat1 <- lat1 * pi / 180
  long2 <- long2 * pi / 180
  lat2 <- lat2 * pi / 180

  R <- 6371 # Earth mean radius in km

  a <- sin((lat2 - lat1) / 2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1) / 2)^2
  d <- R * 2 * asin(sqrt(a))

  return(round(d, round)) # distance in km
}
```

### Exercise 6

```{r}
dn_lq_ak <- dn_lq_ak %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))
```

### Exercise 7

```{r}
dn_lq_ak_min <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarize(closest_distance = min(distance))
```

### Exercise 8

```{r}
summary(dn_lq_ak_min$closest_distance)
```

```{r}
ggplot(dn_lq_ak_min, aes(x = closest_distance)) +
  geom_histogram(binwidth = 5, fill = "orange", color = "red") +
  labs(title = "Distance to Nearest La Quinta from Denny's - Alaska)", 
       x = "Distance (km)", y = "Count")
```

The distribution of distances between Denny's and the nearest La Quinta in Alaska is left skewed indicating a higher concentration of distances near the maximum value. 

View(dn_lq_ak)

### Exercise 9 - NC

View(dennys)

View(laquinta)

```{r}
dn_nc <- dn %>%
  filter(state == "NC")
nrow(dn_nc)
```

```{r}
lq_nc <- lq %>%
  filter(state == "NC")
nrow(lq_nc)
```

```{r}
num_parings <- nrow(dn_nc) * nrow((lq_nc))
```

```{r}
dn_lq_nc <- full_join(dn_nc, lq_nc,
  by = "state"
)
dn_lq_nc
```

```{r}
dn_lq_nc <- dn_lq_nc %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))
```

```{r}
dn_lq_nc_min <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarize(closest_distance = min(distance))
```

```{r}
summary(dn_lq_nc_min$closest_distance)
```

```{r}
ggplot(dn_lq_nc_min, aes(x = closest_distance)) +
  geom_histogram(binwidth = 5, fill = "pink", color = "purple") +
  labs(title = "Distance to Nearest La Quinta from Denny's - North Carolina)", 
       x = "Distance (km)", y = "Count")
```

### Exercise 10 - Texas

```{r}
dn_tx <- dn %>%
  filter(state == "TX")
nrow(dn_tx)
```

```{r}
lq_tx <- lq %>%
  filter(state == "TX")
nrow(lq_tx)
```

```{r}
num_parings <- nrow(dn_tx) * nrow((lq_tx))
```

```{r}
dn_lq_tx <- full_join(dn_tx, lq_tx,
  by = "state"
)
dn_lq_tx
```

```{r}
dn_lq_tx <- dn_lq_tx %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))
```

```{r}
dn_lq_tx_min <- dn_lq_tx %>%
  group_by(address.x) %>%
  summarize(closest_distance = min(distance))
```

```{r}
summary(dn_lq_tx_min$closest_distance)
```

```{r}
ggplot(dn_lq_tx_min, aes(x = closest_distance)) +
  geom_histogram(binwidth = 5, fill = "green", color = "yellow") +
  labs(title = "Distance to Nearest La Quinta from Denny's - Texas)", 
       x = "Distance (km)", y = "Count")
```

The distribution of distances between Denny's and the nearest La Quinta in Texas is right skewed indicating a higher concentration of distances near the minimum value. 


### Exercise 11 - Georgia

```{r}
dn_ga <- dn %>%
  filter(state == "GA")
nrow(dn_ga)
```

```{r}
lq_ga <- lq %>%
  filter(state == "GA")
nrow(lq_ga)
```

```{r}
num_parings <- nrow(dn_ga) * nrow((lq_ga))
```

```{r}
dn_lq_ga <- full_join(dn_ga, lq_ga,
  by = "state"
)
dn_lq_ga
```

```{r}
dn_lq_ga <- dn_lq_ga %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))
```

```{r}
dn_lq_ga_min <- dn_lq_ga %>%
  group_by(address.x) %>%
  summarize(closest_distance = min(distance))
```

```{r}
summary(dn_lq_ga_min$closest_distance)
```

```{r}
ggplot(dn_lq_ga_min, aes(x = closest_distance)) +
  geom_histogram(binwidth = 5, fill = "orange", color = "pink") +
  labs(title = "Distance to Nearest La Quinta from Denny's - Georgia)", 
       x = "Distance (km)", y = "Count")
```

The distribution of distances between Denny's and the nearest La Quinta in Georgia is right skewed indicating a higher concentration of distances near the minimum value. 


### Exercise 12

Mitch Hedberg’s joke held true for the state of Texas because they have the greatest amount of Dennys and La Quintas with a short distance between them.
