---
title: "Lab 05 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Shatavia Bellmon"
date: "2/23/26"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(readr) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
dennys <- read_csv("data/dennys.csv")
laquinta <- read_csv("data/laquinta.csv")
```

### Exercise 1

```{r}
dn <- dennys
```

```{r}
dn_ak <- dn %>%
  filter(state == "AK")
nrow(dn_ak)
```

There are three Denny's in Alaska.

```{r}
lq <- laquinta
```

```{r}
lq_ak <- lq %>%
  filter(state == "AK")
nrow(lq_ak)
```

There are two Laquintas in Alaska.

### Exercise 2

```{r}
num_parings <- nrow(dn_ak) * nrow((lq_ak))
```

There needs to be 6 parings.

### Exercise 3

```{r}
dn_lq_ak <- full_join(dn_ak, lq_ak,
  by = "state"
)
dn_lq_ak
```


### Exercise 4

View(dn_lq_ak)

There are six observations in the dn_lq_ak data frame. The variables in this data frame are address.x, city.x, state, zip.x, longitude.x, latitude.x, address.y, city.y, zip.y, longitude.y, and latitude.y.

### Exercise 5

```{r}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 <- long1 * pi / 180
  lat1 <- lat1 * pi / 180
  long2 <- long2 * pi / 180
  lat2 <- lat2 * pi / 180

  R <- 6371 # Earth mean radius in km

  a <- sin((lat2 - lat1) / 2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1) / 2)^2
  d <- R * 2 * asin(sqrt(a))

  return(round(d, round)) # distance in km
}
```

### Exercise 6

```{r}
dn_lq_ak <- dn_lq_ak %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, longitude.y))
```

### Exercise 7

```{r}
dn_lq_ak_min <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarize(closest_distance = min(distance))
```

### Exercise 8

```{r}
summary(dn_lq_ak_min$closest_distance)
```

```{r}
ggplot(dn_lq_ak_min, aes(x = closest_distance)) +
  geom_histogram(binwidth = 5, fill = "orange", color = "red") +
  labs(title = "Distance to Nearest La Quinta from Denny's - Alaska)", 
       x = "Distance (km)", y = "Count")
```

The distribution of distances between Denny's and the nearest La Quinta in Alaska is left skewed with a greater distance being more prevalent amoung the different locations, 

View(dn_lq_ak)

### Exercise 9


### Exercise 10


### Exercise 11


